Practical Parametricity for GADTs

### 2020-07-15 20-gadts/inf-rules.tex
added IEL section and started supp.tex

@@

If $\mcA$ and $\mcC$ are $\omega_1$-cocomplete, then a functor $F :
\mcA \to \mcC$ is {\em $\omega_1$-cocontinuous} if it preserves
$\omega_1$-directed colimits. We write $[\A,\C]_{\omega_1}$ for 
the category of $\omega_1$-cocontinuous functors from $\A$ to $\C$,
and $\C^\A$ for the category of {\em all} functors from $\A$ to $\C$.
Since (co)limits of functors are computed pointwise, $C^\A$ has all
(co)limits that $\C$ has, and (co)limits of (co)continuous functors
are again (co)continuous. It follows that $[\A,\C]_{\omega_1}$ is
($\omega_1$-)(co)complete whenever $\C$ is.

@@

Now, for any $k$, any $A : \oCPO$, and any $R : \oCPOR(A, B)$, let
$K^\oCPO_A$ be the constantly $A$-valued functor from $\oCPO^k$ to
$\oCPO$ and $K^\oCPOR_R$ be the constantly $R$-valued functor from
$\oCPOR^k$ to $\oCPOR$.  Also let $0$ denote either the initial object
of $\oCPO$ or the initial object of $\oCPOR$, as appropriate.
Observing that, for every $k$, $K^\oCPO_0$ is initial in
$[\oCPO^k,\oCPO]_{\omega_1}$, and $K^\oCPOR_0$ is initial in
$[\oCPOR^k,\oCPOR]_{\omega_1}$, we have that, for each $k$, $K_0 =
(K^\oCPO_0,K^\oCPO_0,K^\oCPOR_0)$ is initial in $\oCPORT_k$. Thus, if
$T = (T^1,T^2,T^*) : \oCPORT_k \to \oCPORT_k$ is an endofunctor on
$\oCPORT_k$ then we can define the $\oCPO$ relation transformer
$\mu T$ to be $\colim{n < \omega_1}{T^n K_0}$.
It is not hard to see that $\mu T$ is given explicitly as 
\begin{equation}\label{eq:mu}
  \mu T = (\mu T^1,\mu T^2,
\colim{n < \omega_1}{(T^nK_0)^*})
\end{equation}
and that, as our notation suggests, it really is a fixpoint for $T$ if
$T$ is $\omega_1$-cocontinuous:
\begin{lemma}\label{lem:fp}
For any $T : [\oCPORT_k,\oCPORT_k]_{\omega_1}$, $\mu T \cong T(\mu T)$.
\end{lemma}
\noindent
The isomorphism is given by the morphisms $(\mathit{in}_1,
\mathit{in}_2) : T(\mu T) \to \mu T$ and $(in_1^{-1}, in_2^{-1}) : \mu
T \to T(\mu T)$ in $\oCPORT_k$. The latter is always a morphism in $\oCPORT_k$,
but the former need not be if $T$ is not $\omega_1$-cocontinuous.

@@

>>>>

\new{Locally presentable categories are the closest approximation to
  small complete categories. It's remarkable that local presentability
  is sufficient not just to give semantics to GADTs, but is also all
  that is needed to ensure that the model induced by that semantics is
  actually parametric.}

@@

If $\mcA$ and $\mcC$ are $\omega_1$-cocomplete, then a functor $F :
\mcA \to \mcC$ is {\em $\omega_1$-cocontinuous} if it preserves
$\omega_1$-directed colimits. We write $[\A,\C]_{\omega_1}$ for the
category of $\omega_1$-cocontinuous functors from $\A$ to $\C$, and
$\C^\A$ for the category of {\em all} functors from $\A$ to $\C$.
Since (co)limits of functors are computed pointwise, $C^\A$ has all
(co)limits that $\C$ has, and (co)limits of (co)continuous functors
are again (co)continuous. It follows that $[\A,\C]_{(\omega_1)}$ is
($\omega_1$-)(co)complete whenever $\C$ is.  {\color{red} In
  particular, if $F : [\C, \C]_{(\omega_1)}$ for a(n)
  ($\omega_1$-)cocomplete category $\C$, then we can define $F^0 =
  \Id$, $F^{\alpha + 1} = F \circ F^\alpha$ at successor ordinals, and
  $F^{\bigcup_{\beta < \alpha} \beta} = \colim{\beta < \alpha}
  F^\beta$ at limit ordinals.}

@@

Now, for any $k$, any $A : \oCPO$, and any $R : \oCPOR(A, B)$, let
$K^\oCPO_A$ be the constantly $A$-valued functor from $\oCPO^k$ to
$\oCPO$ and $K^\oCPOR_R$ be the constantly $R$-valued functor from
$\oCPOR^k$ to $\oCPOR$.  Also let $0$ denote either the initial object
of $\oCPO$ or the initial object of $\oCPOR$, as appropriate.
Observing that, for every $k$, $K^\oCPO_0$ is initial in
$[\oCPO^k,\oCPO]_{\omega_1}$, and $K^\oCPOR_0$ is initial in
$[\oCPOR^k,\oCPOR]_{\omega_1}$, we have that, for each $k$, $K_0 =
(K^\oCPO_0,K^\oCPO_0,K^\oCPOR_0)$ is initial in $\oCPORT_k$. Thus, if
$T = (T^1,T^2,T^*) : \oCPORT_k \to \oCPORT_k$ is an endofunctor on
$\oCPORT_k$ then we can define the $\oCPO$ relation transformer $\mu
T$ to be $\colim{i < \omega_1}{T^i K_0}$. It is not hard to see that
$\mu T$ is given explicitly as
\begin{equation}\label{eq:mu}
  \mu T = (\mu T^1,\mu T^2,
\colim{i < \omega_1}{(T^iK_0)^*})
\end{equation}
and that, as our notation suggests, it really is a fixpoint for $T$ if
$T$ is $\omega_1$-cocontinuous:
\begin{lemma}\label{lem:fp}
For any $T : [\oCPORT_k,\oCPORT_k]_{\omega_1}$, $\mu T \cong T(\mu T)$.
\end{lemma}
\noindent
The isomorphism is given by the morphisms $(\mathit{in}_1,
\mathit{in}_2) : T(\mu T) \to \mu T$ and $(in_1^{-1}, in_2^{-1}) : \mu
T \to T(\mu T)$ in $\oCPORT_k$. The latter is always a morphism in $\oCPORT_k$,
but the former need not be if $T$ is not $\omega_1$-cocontinuous.

@@

\section{The Identity Extension Lemma}\label{sec:iel}

The standard definition of the graph for a morphism $f : A \to B$ in
$\set$ is the relation $\graph{f} : \rel(A,B)$ defined by $(x,y) \in
\graph{f}$ iff $fx = y$. This definition naturally generalizes to
associate to each natural transformation between $k$-ary functors on
$\oCPO$ a $k$-ary $\oCPO$ relation transformer as follows.

First, let $f : A \to B$ be a morphism in $\oCPO$. Then the graph of
the underlying function $|f| : |A| \to |B|$ in $\set$ is an $\oCPO$
relation.  Indeed, if $(a_i, b_i) : \graph{|f|}$ for all $i < \omega$
is a chain, then $|f| a_i = b_i$ for all $i$, and consequently
$|f|(\bigvee_{i < \omega} a_i) = \bigvee_{i < \omega} (|f| a_i) =
\bigvee_{i < \omega} b_i$, i.e., $(\bigvee_{i < \omega} a_i,
\bigvee_{i < \omega} b_i) : \graph{|f|}$. We therefore define the {\em
  $\oCPO$ graph of $f$} in $\oCPO$ to be $\graph{f} =
\graph{|f|}$. Note in particular that if $A : \oCPO$ then
$\graph{\id_A}$ is an $\oCPO$ relation. We denote this $\oCPO$
relation, called the {\em equality relation on $A$}, by $\Eq_A$. It
coincides exactly with the equality relation on the underlying set
$|A|$ of $A$.

The notion of an $\oCPO$ graph relation naturally generalizes to
associate to each natural transformation between $k$-ary functors on
$\oCPO$ a $k$-ary $\oCPO$ relation transformer as follows. Recall
that, since $\oCPO$ is a locally $\omega_1$-presentable category,
Proposition 1.6.1 of~\cite{ar94} ensures that it has a (strong epi,
mono) factorization system. We then have:

\begin{dfn}\label{dfn:graph-nat-transf}
If $F, G: \oCPO^k \to \oCPO$ and $\alpha : F \to G$ is a natural
transformation, then the functor $\graph{\alpha}^*: \oCPOR^k \to
\oCPOR$ is defined as follows. Given $R_1 : \oCPOR(A_1, B_1),...,R_k :
\oCPOR(A_k,B_k)$, let $\iota_{R_i} : R_i \hookrightarrow A_i \times
B_i$, for $i = 1,...,k$, be the inclusion of $R_i$ as a
sub$\oCPO$ of $A_i \times B_i$, let $h_{\overline{A \times B}}$
be the unique morphism making the diagram
{\footnotesize\[\begin{tikzcd}[row sep = large] F\overline{A}
  &F(\overline{A \times B}) \ar[l, "{F\overline{\pi_1}}"'] \ar[r,
    "{F\overline{\pi_2}}"] \ar[d, dashed, "{h_{\overline{A \times
          B}}}"] &F\overline{B} \ar[r, "{\alpha_{\ol{B}}}"]
  &G\overline{B} \\ &F\overline{A} \times G\overline{B} \ar[ul,
    "{\pi_1}"] \ar[urr, "{\pi_2}"']
\end{tikzcd}\]}
\noindent
\!\!commute, and let $h_{\overline{R}} : F\overline{R} \to
F\overline{A} \times G\overline{B}$ be $h_{\overline{A \times B}}
\circ F\overline{\iota_R}$. Further, let $\alpha^\wedge\overline{R}$
be the
%If $f = e \circ m = e' \circ m'$ then $cod e = dom m$ is the same as
%$cod e' = dom m'$ up to isomorphism. This follows easily from the
%definition of a factorization system.
subobject through which $h_{\overline{R}}$ is factorized by the
(strong epi, mono) factorization system of $\oCPO$, as shown in the
following diagram: {\footnotesize\[\begin{tikzcd} F\overline{R}
  \ar[rr, "{h_{\overline{R}}}"] \ar[dr, twoheadrightarrow,
    "{q_{\alpha^\wedge\overline{R}}}"'] &&F\overline{A} \times
  G\overline{B} \\ &\alpha^\wedge\overline{R} \ar[ur, hookrightarrow,
    "{\iota_{\alpha^\wedge\overline{R}}}"']
\end{tikzcd}\]}

\noindent
Then $\alpha^\wedge\overline{R} : \oCPOR(F\overline{A},
G\overline{B})$ by construction, so the action of $\graph{\alpha}^*$
on objects of $\oCPOR$ can be given by $\langle \alpha \rangle^*
\overline{(A,B,R)} = (F\overline{A}, G\overline{B},
\iota_{\alpha^\wedge \overline{R}}\alpha^\wedge\overline{R})$. The
action of $\graph{\alpha}^*$ on morphisms of $\oCPOR$ is given by
$\graph{\alpha}^*\overline{(\beta, \beta')} = (F\overline\beta,
G\overline\beta')$.
\end{dfn}

The next lemma shows that the data in
Definition~\ref{dfn:graph-nat-transf} actually yield a relation
transformer $\graph{\alpha} = (F, G, \graph{\alpha}^*)$. We call this
the {\em $\oCPO$ graph relation transformer} for $\alpha$.

\begin{lemma}\label{lem:graph-reln-functors}
If $F,G : [\oCPO^k,\oCPO]_{\omega_1}$, and if $\alpha : F \to G$ is a natural
transformation, then $\graph{\alpha}$ is in $\oCPORT_k$.
\end{lemma}
\begin{proof}
Clearly, $\graph{\alpha}^*$ is $\omega_1$-cocontinuous, so
$\graph{\alpha}^* : [\oCPOR^k,\oCPOR]_{\omega_1}$. Now, suppose
$\overline{R : \oCPOR(A, B)}$, $\overline{S : \oCPOR(C, D)}$, and
$\overline{(\beta, \beta') : R \to S}$. We want to show that there
exists a morphism $\epsilon : \alpha^\wedge\overline{R} \to
\alpha^\wedge\overline{S}$ such that the diagram on the left below
commutes. Since $\ol{(\beta,\beta') : R \to S}$, there exist
$\overline{\gamma : R \to S}$ such that each diagram in the middle
commutes.  Moreover, since both $h_{\overline{C \times D}} \circ
F(\overline{\beta \times \beta'})$ and $(F\overline{\beta} \times
G\overline{\beta'}) \circ h_{\overline{A \times B}}$ make the diagram
on the right commute, they must be equal.
\begin{figure*}[ht]
\vspace*{-0.1in}
  \hspace*{-0.85in}
  \begin{minipage}[b]{0.25\linewidth}
 {\small    \[
    \begin{tikzcd}
        \alpha^\wedge\overline{R}
        \ar[r, hookrightarrow, "{\iota_{\alpha^\wedge\overline{R}}}"]
        \ar[d, "{\epsilon}"']
        & F\overline{A} \times G\overline{B}
        \ar[d, "{F\overline{\beta} \times G\overline{\beta'}}"] \\
        \alpha^\wedge\overline{S}
        \ar[r, hookrightarrow, "{\iota_{\alpha^\wedge\overline{S}}}"']
        & F\overline{C} \times G\overline{D}
    \end{tikzcd}
    \]}
\end{minipage}
\begin{minipage}[b]{0.25\linewidth}
{\small    \[
    \begin{tikzcd}
        R_i
        \ar[d, "{\gamma_i}"']
        \ar[r, hookrightarrow, "{\iota_{R_i}}"]
        &A_i \times B_i
        \ar[d, "{\beta_i \times \beta'_i}"] \\
        S_i
        \ar[r, hookrightarrow, "{\iota_{S_i}}"]
        &C_i \times D_i
    \end{tikzcd}
    \]}
\end{minipage}
\begin{minipage}[b]{0.25\linewidth}
{\footnotesize \[
      \begin{tikzcd}[row sep = large]
          F\overline{C}
          &F\overline{C} \times F\overline{D}
          \ar[l, "{\pi_1}"'] \ar[r, "{\pi_2}"]
          &F\overline{D}
          \ar[r, "{\alpha_{\ol{D}}}"]
          &G\overline{D}\\
          &F(\overline{A \times B})
          \ar[u, dashed, "{\exists !}"]
          \ar[ul, "{F\ol{\pi_1} \circ F(\overline{\beta \times \beta'})}"]
          \ar[urr, "{\alpha_{\ol{D}} \circ F\ol{\pi_2} \circ F(\overline{\beta \times \beta'})}"']
      \end{tikzcd}
      \]}
\end{minipage}
\end{figure*}
We therefore get that the right-hand square in the diagram on the left
below commutes, and thus that the entire diagram does as well.
Finally, by the left-lifting property of
$q_{\alpha^\wedge\overline{R}}$ with respect to
$\iota_{\alpha^\wedge\overline{S}}$ given by the (strong epi, mono)
factorization system there exists an $\epsilon$ such that the diagram
on the right below commutes.
      \begin{tikzcd}
          F\overline{R}
          \ar[d, "{F\overline{\gamma}}"']
          \ar[r, hookrightarrow, "{F\overline{\iota_R}}"]
          \ar[rr, bend left, "{h_{\overline{R}}}"]
          &F(\overline{A \times B})
          \ar[d, "{F(\overline{\beta \times \beta'})}"]
          \ar[r, "{h_{\overline{A \times B}}}"]
          &F\overline{A} \times G\overline{B}
          \ar[d, "{F\overline{\beta} \times F\overline{\beta'}}"] \\
          F\overline{S}
          \ar[r, hookrightarrow, "{F\overline{\iota_S}}"']
          \ar[rr, bend right, "{h_{\overline{S}}}"']
          &F(\overline{C \times D})
          \ar[r, "{h_{\overline{C \times D}}}"']
          &F\overline{C} \times G\overline{D}
      \end{tikzcd}
\qquad
      \begin{tikzcd}
          F\overline{R}
          \ar[d, "{F\overline{\gamma}}"']
          \ar[r, twoheadrightarrow, "{q_{\alpha^\wedge\overline{R}}}"]
          &\alpha^\wedge\overline{R}
          \ar[d, dashed, "{\epsilon}"]
          \ar[r, hookrightarrow, "{\iota_{\alpha^\wedge\overline{R}}}"]
          &F\overline{A} \times G\overline{B}
          \ar[d, "{F\overline{\beta} \times G\overline{\beta'}}"] \\
          F\overline{S}
          \ar[r, twoheadrightarrow, "{q_{\alpha^\wedge\overline{S}}}"']
          &\alpha^\wedge\overline{S}
          \ar[r, hookrightarrow, "{\iota_{\alpha^\wedge\overline{S}}}"']
          &F\overline{C} \times G\overline{D}
      \end{tikzcd}
\end{proof}

If $f : A \to B$ is a morphism in $\oCPO$ then the definition of the
$\oCPO$ graph relation transformer $\graph{f}$ for $f$ as a natural
transformation between $0$-ary functors $A$ and $B$ coincides with the
definition of $\graph{f}$ for $f$ as a morphism in $\oCPO$ given in
the second paragraph of this section.  As a result, $\oCPO$ graph
relation transformers are a reasonable extension of $\oCPO$ graph
relations to functors.

To prove the IEL, we will need to know that the equality $\oCPO$
relation transformer preserves equality relations in $\oCPOR$; see
Equation~\ref{eq:eq-pres-eq} below. This will follow from the next
lemma, which shows how to compute the action of an $\oCPO$ graph
relation transformer on any $\oCPO$ graph relation.

\begin{lemma}\label{lem:eq-reln-equalities}
If $\alpha : F \to G$ is a morphism in $[\oCPO^k, \oCPO]_{\omega_1}$
and $f_1: A_1 \to B_1, ..., f_k : A_k \to B_k$, then $\graph{\alpha}^*
\graph{\overline{f}} = \graph{G \ol{f} \circ \alpha_{\ol{A}}} =
\graph{\alpha_{\ol{B}} \circ F \ol{f}}$.
\end{lemma}
\begin{proof}
Since $h_{\overline{A \times B}}$ is the unique morphism making the
bottom triangle of the diagram on the left below commute, and since
$h_{\graph{\overline{f}}} = h_{\overline{A \times B}} \circ F
\,\ol{\iota_{\graph{f}}} = h_{\overline{A \times B}} \circ F
\overline{\graph{\id_A, f}}$, the universal property of the
product depicted in the diagram on the right gives
$h_{\graph{\overline{f}}} = \graph{\id_{F \ol{A}}, \alpha_{\ol{B}}
\circ F\ol{f}} : F \ol{A} \to F \ol{A} \times G \ol{B}$.
\begin{figure*}[ht]
  \vspace*{-0.15in}
  \begin{minipage}[b]{0.45\linewidth}
{\footnotesize
\[\begin{tikzcd}[row sep = large]
        &F\overline{A}
        \ar[d, "{F \overline{\graph{\id_A, f}}}" description]
        \ar[dl, equal]
        \ar[dr, "{F\ol{f}}"]\\
        F\overline{A}
        &F(\overline{A \times B})
        \ar[l, "{F\overline{\pi_1}}"']
        \ar[r, "{F\overline{\pi_2}}"]
        \ar[d, "{h_{\overline{A \times B}}}"]
        &F\overline{B}
        \ar[r, "{\alpha_{\ol{B}}}"]
        &G\overline{B}\\
        &F\overline{A} \times G\overline{B}
        \ar[ul, "{\pi_1}"] \ar[urr, "{\pi_2}"']
\end{tikzcd}\]}
\end{minipage}
  \begin{minipage}[b]{0.45\linewidth}
{\footnotesize
\[
      \begin{tikzcd}[row sep = large]
          F\overline{A}
          &F\overline{A} \times G\overline{B}
          \ar[l, "{\pi_1}"'] \ar[r, "{\pi_2}"]
          &G\overline{B}\\
          &F\overline{A}
          \ar[u, dashed, "{\exists !}"]
          \ar[ul, equal]
          \ar[r, "{F\ol{f}}"']
          &F{\ol{B}}
          \ar[u, "{\alpha_{\ol{B}}}"']
      \end{tikzcd}
      \]}
\end{minipage}
\end{figure*}

\vspace*{-0.1in}

\noindent
Moreover, $\graph{\id_{F \ol{A}}, \alpha_{\ol{B}} \circ F\ol{f}}$ is a
monomorphism in $\oCPO$ because $\id_{F \ol{A}}$ is, so its (strong
epi, mono) factorization gives $\iota_{\alpha^\wedge
  \graph{\overline{f}}} = \langle \id_{F \ol{A}}, \alpha_{\ol{B}}
\circ F\ol{f} \rangle$, and thus that $\alpha^\wedge
\graph{\overline{f}} = F\overline{A}$.
%If $e \circ m = e \circ m'$ then $m = m'$ is trivial from the
%definition of a factorization system.
Then $\iota_{\alpha^\wedge \graph{\overline{f}}} \alpha^\wedge
\graph{\overline{f}} = \graph{\id_{F \ol{A}}, \alpha_{\ol{B}} \circ
  F\ol{f}} (F \ol{A}) = \graph{ \alpha_{\ol{B}} \circ F\ol{f}
}^*$. Then $\graph{ \alpha }^* \graph{ \overline{f} } =
(F\overline{A}, G\overline{B}, \iota_{\alpha^\wedge
  \graph{\overline{f}}}\, \alpha^\wedge \graph{\overline{f}}) =
(F\overline{A}, G\overline{B}, \graph{ \alpha_{\ol{B}} \circ F\ol{f}
}^*) = \graph{ \alpha_{\ol{B}} \circ F\ol{f} }$, and, finally,
$\alpha_{\ol{B}} \circ F\ol{f} = G\ol{f} \circ \alpha_{\ol{A}}$ by
naturality of $\alpha$.
\end{proof}

The {\em equality $\oCPO$ relation transformer} on $F :
[\oCPO^k,\oCPO]_{\omega_1}$ is defined to be $\Eq_F =
\graph{\id_{F}}$. Specifically, $\Eq_F = (F, F, \Eq_F^*)$ with
$\Eq_F^* = \graph{\id_{F}}^*$, and Lemma~\ref{lem:eq-reln-equalities}
indeed ensures that
\begin{equation}\label{eq:eq-pres-eq}
\Eq^*_F \ol{\Eq_A}
= \graph{\id_F}^* \graph{\id_{\ol{A}}}
= \graph{F \id_{\ol{A}} \circ (\id_F)_{\ol{A}}}
= \graph{\id_{F\ol{A}} \circ \id_{F\ol{A}}}
= \graph{\id_{F\ol{A}}}
= \Eq_{F\ol{A}}
\end{equation}
for all $\ol{A : \oCPO}$.  Graph $\oCPO$ relation transformers in
general, and equality $\oCPO$ relation transformers in particular,
extend to $\oCPO$ relation environments in the obvious ways.  Indeed,
if $\rho, \rho' : \oCPOenv$ and $f : \rho \to \rho'$, then the {\em
  graph $\oCPO$ relation environment} $\graph{f}$ is defined pointwise
by $\graph{f} \phi = \graph{f \phi}$ for every $\phi$, which entails
that $\pi_1 \graph{f} = \rho$ and $\pi_2 \graph{f} = \rho'$. In
particular, the {\em equality $\oCPO$ relation environment} $\Eq_\rho$
is defined to be $\graph{\id_{\rho}}$, which entails that $\Eq_\rho
\phi = \Eq_{\rho \phi}$ for every $\phi$.  With these definitions in
hand, we can state and prove both an Identity Extension Lemma and a
Graph Lemma for our calculus.

\begin{thm}[IEL]\label{thm:iel}
  If $\rho : \oCPOenv$ and $\Gamma; \Phi \vdash F$ then
  $\oCPORsem{\Gamma; \Phi \vdash F} \Eq_\rho = \Eq_{\oCPOsem{\Gamma;
      \Phi \vdash F}\rho}$.
\end{thm}
The proof is by induction on the structure of $F$. Only the $\Nat$,
application, and fixpoint cases are non-routine. The latter two use
Lemma~\ref{lem:eq-reln-equalities}. The fixpoint case also uses the
observation that, for every $i < \omega_1$, the following intermediate
results can be proved by simultaneous induction with
Theorem~\ref{thm:iel}: for any $H$, $\rho$, $A$, and any subformula
$J$ of $H$, both
$T^i_{H,\Eq_{\rho}} K_0\, \ol{\Eq_A}
= (\Eq_{(T^\set_{H,\rho})^i K_0})^*
\ol{\Eq_A}$\;
and\;
\[\begin{array}{ll}
&  \oCPORsem{\Gamma; \Phi, \phi, \ol{\alpha} \vdash J} \Eq_{\rho} [\phi
  := T^i_{H,\Eq_{\rho}} K_0] \overline{[\alpha := \Eq_A]}\\
=  &
\oCPORsem{\Gamma; \Phi, \phi,
  \ol{\alpha} \vdash J} \Eq_{\rho} [\phi := \Eq_{(T^\oCPO_{H,\rho})^i K_0}]
\overline{[\alpha := \Eq_A]}
\end{array}\]
hold. With these results in hand, the proof follows easily. It is
given in detail in the accompanying anonymous supplementary
material. As noted there, if functorial variables of arity greater
than $0$ were allowed to appear in the bodies of $\mu$-types, then the
IEL would fail.

\begin{lemma}[Graph Lemma]\label{lem:graph}
If $\rho, \rho' : \oCPOenv$, $f : \rho \to \rho'$, and $\Gamma; \Phi
\vdash F$, then $\graph{\oCPOsem{\Gamma; \Phi \vdash F} f} =
\oCPORsem{\Gamma; \Phi \vdash F}\graph{f}$.
\end{lemma}
\begin{proof}
Applying Lemma~\ref{lem:rel-transf-morph} to the morphisms $(f,
\id_{\rho'}) : \graph{f} \to \Eq_{\rho'}$ and $(\id_{\rho}, f) :
\Eq_{\rho} \to \graph{f}$ of relation environments gives that
\begin{multline*}
(\oCPOsem{\Gamma; \Phi \vdash F}f, \oCPOsem{\Gamma; \Phi \vdash F}\id_{\rho'})
  = \oCPORsem{\Gamma; \Phi \vdash F} (f, \id_{\rho'}) \\
:\oCPORsem{\Gamma; \Phi \vdash F}\graph{f} \to 
  \oCPORsem{\Gamma; \Phi \vdash F}\Eq_{\rho'}
\end{multline*}
and
\begin{multline*}
(\oCPOsem{\Gamma; \Phi \vdash F}\id_{\rho}, \oCPOsem{\Gamma; \Phi \vdash F}f)
  = \oCPORsem{\Gamma; \Phi \vdash F} (\id_{\rho}, f) \\
: \oCPORsem{\Gamma; \Phi \vdash F}\Eq_{\rho} 
  \to \oCPORsem{\Gamma; \Phi \vdash F}\graph{f}
\end{multline*}
Expanding the first equation gives that if $(x,y) \in \oCPORsem{\Gamma;
  \Phi \vdash F}\graph{f}$ then \[(\oCPOsem{\Gamma; \Phi \vdash F} f\,
x, \oCPOsem{\Gamma; \Phi \vdash F}\id_{\rho'}\, y) \in \oCPORsem{\Gamma;
  \Phi \vdash F}\Eq_{\rho'}\] So $\oCPOsem{\Gamma; \Phi \vdash
  F}\id_{\rho'}\, y = \id_{\oCPOsem{\Gamma; \Phi \vdash F}\rho'}\, y =
y$ and $\oCPORsem{\Gamma; \Phi \vdash F}\Eq_{\rho'} =
\Eq_{\oCPOsem{\Gamma; \Phi \vdash F}\rho'}$, and if $(x,y) \in
\oCPORsem{\Gamma; \Phi \vdash F}\graph{f}$ then $(\oCPOsem{\Gamma; \Phi
  \vdash F} f\, x, y) \in \Eq_{\oCPOsem{\Gamma; \Phi \vdash F}\rho'}$,
i.e., $\oCPOsem{\Gamma; \Phi \vdash F} f\, x = y$, i.e., $(x, y) \in
\graph{\oCPOsem{\Gamma; \Phi \vdash F} f}$.  So, we have that
$\oCPORsem{\Gamma; \Phi \vdash F}\graph{f} \subseteq
\graph{\oCPOsem{\Gamma; \Phi \vdash F}f}$.  Expanding the second
equation gives that if $x \in \oCPOsem{\Gamma; \Phi \vdash F}\rho$ then
$(\oCPOsem{\Gamma; \Phi \vdash F}\id_{\rho}\, x,$\\
$\oCPOsem{\Gamma; \Phi
  \vdash F} f\, x) \in \oCPORsem{\Gamma; \Phi \vdash F}\graph{f}$.  Then
$\oCPOsem{\Gamma; \Phi \vdash F}\id_{\rho}\, x = \id_{\oCPOsem{\Gamma;
    \Phi \vdash F}\rho} x = x$, so for any $x \in \oCPOsem{\Gamma; \Phi
  \vdash F}\rho$ we have that $(x, \oCPOsem{\Gamma; \Phi \vdash F}f\,
x) \in \oCPORsem{\Gamma; \Phi \vdash F}\graph{f}$.  Moreover, $x \in
\oCPOsem{\Gamma; \Phi \vdash F}\rho$ if and only if $(x,
\oCPOsem{\Gamma; \Phi \vdash F} f\, x) \in \graph{\oCPOsem{\Gamma; \Phi
    \vdash F}f}$ and, if $x \in \oCPOsem{\Gamma; \Phi \vdash F}\rho$
then $(x, \oCPOsem{\Gamma; \Phi \vdash F} f\, x) \in \oCPORsem{\Gamma;
  \Phi \vdash F} \graph{f}$, so if $(x, \oCPOsem{\Gamma; \Phi \vdash F}
f\, x) \in \graph{\oCPOsem{\Gamma; \Phi \vdash F}f}$ then $(x,
\oCPOsem{\Gamma; \Phi \vdash F} f\, x) \in \oCPORsem{\Gamma; \Phi \vdash
  F} \graph{f}$, i.e., $\graph{\oCPOsem{\Gamma; \Phi \vdash F}f}
\subseteq \oCPORsem{\Gamma; \Phi \vdash F} \graph{f}$.
\end{proof}
